A Deterministic and Terminating Semantics for the Synchronous Programming Language~\CEU

### 2018-02-18 06-appendix.tex
Adiciona lema sobre irredutibilidade NST ao apêndice
I.e., "case", "subcase", e "subsubcase".

\makeatletter
\newtheoremstyle{subcase}%
  {\thm@preskip\topsep \divide\thm@preskip\tw@}% space above
  {\thm@postskip\thm@preskip}% space below
  {\addtolength{\@totalleftmargin}{\parindent}%
    \addtolength{\linewidth}{-\parindent}%
    \parshape 1 \parindent \linewidth
    \normalfont}% body font
  {\z@}% indent amount
  {\itshape}% head font
  {.}% punctuation after head
  {.5em}% spacing after head
  {}% head spec
\newtheoremstyle{subsubcase}%
  {\thm@preskip\topsep \divide\thm@preskip\tw@}% space above
  {\thm@postskip\thm@preskip}% space below
  {\addtolength{\@totalleftmargin}{\parindent}%
    \addtolength{\linewidth}{-\parindent}%
    \parshape 1 2\parindent \linewidth
    \normalfont}% body font
  {\z@}% indent amount
  {\itshape}% head font
  {.}% punctuation after head
  {.5em}% spacing after head
  {}% head spec
\makeatletter

\theoremstyle{remark}
\newtheorem{case}{Case}

\theoremstyle{subcase}
\newtheorem{subcase}{Case}
\numberwithin{subcase}{case}

\theoremstyle{subsubcase}
\newtheorem{subsubcase}{Case}
\numberwithin{subsubcase}{subcase}

\AtBeginEnvironment{proof}{\setcounter{case}{0}}
\AtBeginEnvironment{case}{\setcounter{subcase}{0}}
\AtBeginEnvironment{subcase}{\setcounter{subsubcase}{0}}

@@

\begin{theorem}\label{thm.det-out-pop-n}
  If~$\delta\out[n]\delta_1$ and~$\delta\out[n]\delta_2$
  then~$\delta_1=\delta_2$.
\end{theorem}
\begin{proof}
  By induction on~$n$.
  %%
  The theorem is trivially true for~$n=0$ and follows directly from
  Lemma~\ref{lem.det-out} for~$n=1$.  (Note that, for~$n>1$, the theorem is
  vacuously true for~$\outpush$ transitions.  By the format of
  rules~\R{push} and~\R{pop}, transitions~$\outpush$ cannot be applied more
  than once in a row and cannot occur after a~$\outpop$ transition.)

@@

>>>>

@@

\begin{theorem}\label{thm.det-out-pop-n}
  If~$\delta\out[n]\delta_1$ and~$\delta\out[n]\delta_2$
  then~$\delta_1=\delta_2$.
\end{theorem}
\begin{proof}
  By induction on~$n$.
  %%
  The theorem is trivially true for~$n=0$ and follows directly from
  Lemma~\ref{lem.det-out} for~$n=1$.  (Note that for~$n>1$ the theorem is
  vacuously true for~$\outpush$ transitions---by the format of
  rules~\R{push} and~\R{pop}, transitions~$\outpush$ cannot be applied more
  than once in a row and cannot occur after a~$\outpop$ transition.)

@@

\begin{definition}\label{def.Hnst}
  A description~$\delta=\<p,n,e>$ is \emph{nested-irre\-ducible}
  iff~$p=\ceu{\Nop},\ceu{\Break}$ or~$\isblocked(p,n)$~is true.
  Nested-irreducible descriptions serve as normal forms for~$\nst$
  transitions: their program~$p$ and stack-level~$n$ embody the result of an
  exhaustive number of~$\nst$ transitions.  We will write~$\delta_\Hnst$ to
  indicate that description~$\delta$ is nested-irreducible.
\end{definition}

The next lemma justifies the use of qualifier ``irreducible'' in
Definition~\ref{def.Hnst}.

\begin{lemma}\label{lem.irr-nst-n}
  If~$\delta\nst[n]\delta_\Hnst'$ then, for all~$i\ne{n}$, there is
  no~$\delta_\Hnst''$ such that~$\delta\nst[i]\delta''_\Hnst$.
\end{lemma}
\begin{proof}
  By contradiction.  Let~$\delta\nst[n]\delta'_\Hnst$, for some~$n\ge0$.
  There are two cases.
  \begin{case}\label{lem.irr-nst-n-case1}
    Suppose there are~$i>n$ and~$\delta''_\Hnst$ such
    that~$\delta\nst[i]\delta''$.
    %%
    Then, by definition of~$\nst[i]$,
    \begin{equation}\label{lem.irr-nst-n-eq1}
      \delta\nst[n]\delta'\nst[n+1]\delta_1'\nst[n+2]\cdots\nst[i]\delta''.
    \end{equation}
    Since~$\delta'=\<p',n,e'>$ is nested-irreducible,
    $p=\ceu{\Nop},\ceu{\Break}$ or~$\isblocked(p',n)$.  In either case, by
    the definition of~$\nst$, there is no~$\delta_1'$ such
    that~$\delta'\nst[1]\delta_1'$, which
    contradicts~\eqref{lem.irr-nst-n-eq1}.  Therefore, no such~$i$ can
    exist.
  \end{case}
  \begin{case}
    Suppose there are~$0\le{i}<n$ and~$\delta''_\Hnst$ such
    that~$\delta\nst[i]\delta''$.  Then since~$n>i$, by
    Case~\ref{lem.irr-nst-n-case1}, $\delta'$~could not exist, which is
    absurd.  Therefore, the assumption that there is such~$i$ is
    false.\qedhere
  \end{case}
\end{proof}
